---
title: "RMDdraft1MLMS"
output:
  minidown::mini_document:
    framework: water
    theme: light
    code_folding:
      source: show
      output: show
      message: hide
      warning: hide
      error: show
    toc: true
    toc_float: true
    toc_highlight: true
    tabset: true
    number_sections: true
    anchor_sections: false
    self_contained: false
    code_download: true
    math: "katex"
    keep_md: true
    fig_caption: true
vignette: >
  %\VignetteIndexEntry{Writing Vignette with the 'minidown' Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(results.folding = NULL)
`%>%` <- magrittr::`%>%`
# load R packages
library(readr)
library(isoreader)
library(MLMS)
library(katex)
library(png)
library(grid)
library(DT)
library(pracma)
library(kableExtra)
library(InterpretMSSpectrum)
library(ggplot2)
library(plotly)
library(bookdown)
library(doParallel)
library(splancs)
```

This article introduces the MLMS R library and is a draft of a vignette[@R-base].

The github link for MLMS is 
[github.com/lilyacb/MLMS](https://github.com/lilyacb/MLMS).  Open this document in a browser for the links to work.

# Accessing information and making tables

Tables summarizing .dxf file contents can be produced with *read_summary()*.
```{r,message=F,fig.cap="Summary of file contents"}
  datafile<-"170525_NaHCO3 L + NaCl L_.dxf"
  #data_file<-"https://github.com/lilyacb/MLMS/blob/main/Data/170525_NaHCO3%2#0L%20%2B%20NaCl%20L_.dxf"  # would work if changed to csv!!
  #file.summ<-read_csv(data_file)
  # Print table without kable
  fileI.summ<-read_summary(datafile)
```
<br><br>

You can use kable to make a fancier table.
```{r,message=F}
  #class.source='details hide'
 # Using kable
  knitr::kable(fileI.summ,caption="170525_NaHCO3 L + NaCl L_.dxf file summary")
```
<br><br>

You can print information contained in the file_info, vendor_info and raw_data tabs of the .dxf file.

Get file information with *file_info()*.
```{r,message=FALSE}
# Can get file information
fi.df<-file_info(files=datafile)
knitr::kable(fi.df,caption="170525_NaHCO3 L + NaCl L_.dxf file information")
```
<br>

Get the vendor data table with *vendor_info()*.
```{r,message=F}
# Can get vendor info
vi.df<-vendor_info(datafile)
kbl(head(vi.df)[1:3,],caption="170525_NaHCO3 L + NaCl L_.dxf vendor data") %>% 
  kable_paper() %>%
  scroll_box(width=5,height = "200px")
```
<br>

Get the raw data using *raw_data()*
```{r,message=FALSE}
# Can get the raw data
raw.df<-raw_data(datafile)
knitr::kable(head(raw.df),caption="170525_NaHCO3 L + NaCl L_.dxf raw data")
```
<br>

Get the resistor information using *resistor_data()*
```{r,message=FALSE}
# Can get the resistor information
resist<-resistor_data(datafile)
knitr::kable(resist,caption="170525_NaHCO3 L + NaCl L_.dxf resistor information")
```
<br>

Get the isotopic reference values with *reference_values_ratio()*
```{r,message=FALSE}
# Can get isotopic reference values with ratios
stand_ratio<-reference_values_ratio(datafile)
knitr::kable(stand_ratio,caption="170525_NaHCO3 L + NaCl L_.dxf isotopic reference values with ratios")
```
<br>

Use DT to render larger tables neatly. You can show only a few lines, have a search bar, filters and more.

```{r,message=FALSE}
datatable(vi.df,#filter="top",
          options=list(pageLength=5,scrollX=T))
```
<br><br>

# MS data processing

## Peak areas

Peak areas were calculated using the *trapz* package, which implements numerical integration via the trapezoid rule.

Math with KaTex produces nice equations. You can highlight formulas or text.<br>
<br>
The trapezoid rule:
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">
$$
\int\limits_a^b f(x)dx\approx
\sum_{n=0}^{N-1}{\frac{1}{2}(f_n+f_{n+1})(\varDelta x)_n}
$$
</div>
<br>
You can insert images into an RMD doc.
```{r,fig.cap=paste("Graphical illustration of the trapezoid rule for numerical integration")}
trapImg<-readPNG("trapezoidalRuleImg.png")
grid.raster(trapImg)
```
<br>

Calculate the peak areas using *trap_area_allPks()*.
```{r,message=FALSE}
# Can get peak areas
rawN<-"v44.mV"
areaPks<-trap_area_allPks(raw.df,vi.df,rawN)
knitr::kable(areaPks,caption="170525_NaHCO3 L + NaCl L_.dxf all peak areas")
```
<br><br>


## Data visualization
<style>
p.caption {
  font-size: 0.9em;
  font-style: italic;
  color: grey;
  margin-right: 10%;
  margin-left: 10%;  
  text-align: justify;
}
</style>

You can graph the intensity vs retention time using *plot_ms()*.
```{r,fig.cap="Intensity_all vs Rt for 170525_NaHCO3_L_+_NaCl",message=FALSE}
# Intensity_All vs Rt for an experiment
ia.xname<-"Rt"
ia.yname<-"Intensity_All"
# plot
plot_ms(vi.df,ia.xname,ia.yname)
```
<br><br>


Plot individual peaks in an experiment with *plot_individual_peaks()*.
```{r,message=FALSE,fig.cap=paste("Peak 1 intensity (v44.mV) vs. time")}
# Can plot peaks in an experiment individually (Intensity (mV) vs Rt)
time.s<-as.numeric(raw.df$time.s)
start.v1<-as.numeric(vi.df$Start)
end.v1<-as.numeric(vi.df$End)
v44<-as.numeric(raw.df$v44.mV)
# plot just the first peak to inspect
peak1.p<-plot_individual_peaks(start.v1,end.v1,time.s,v44,"1","v44.mV")
```
<br><br>

Plot the raw data, intensity vs retention time, with *gg_raw_plot()*--make function!!

Use ggplot to plot the raw (**redo colour label in legend**)
```{r,fig.caption="170525_NaHCO3 L + NaCl L_.dxf raw data"}
raw.dat<-read.table("LLrawdat")
ggplot(raw.dat,aes(x=time.s,y=v44.mV))+
  geom_line(aes(color="v44.mV"))+
  geom_line(aes(x=time.s,y=v45.mV,color="v45.mV"))+
  geom_line(aes(x=time.s,y=v46.mV,color="v46.mV"))+
  labs(title="170525_NaHCO3 L + NaCl U",x="time.s",y="v44_v45_v46.mV")
```
<br>

Plot the raw data of all files in a directory and export to a pdf file using *generic_plot_all_raw()* **change to ggplot version**--new func!
```{r,message=FALSE}
# Can plot raw data of all files in a directory
# Get all filenames of .dxf files in the directory
LLdir<-"NaHCO3_L_+_NaCl_L"
fileNames<-all_filenames(LLdir)
setwd(LLdir)
rawList<-raw_data_all(fileNames)
setwd("~/Desktop/EuropaMLMS/rmdMLMS")
# plot all raw data
generic_plot_all_raw(rawList)
```
<br>

## Sorting and quality checks

### Automated sorting of a directory of .dxf files
A directory of .dxf files can be organized by Identifier_1 and Preparation method.
<br><br>

```{r,echo=FALSE,message=FALSE}
sort_by_identifier_1<-function(path){
  all_filenames<-all_filenames(path)
  print(all_filenames)
  un_identifiers<-unique_identifiers(all_filenames)
  print(un_identifiers)
  num_identifiers<-length(un_identifiers)
  path_ID<-paste(path,"/",sep="")
  for(i in seq(1:num_identifiers)){
    # get all files with un_identifier[i]
    I1_files<-identifier_1_files(all_filenames,un_identifiers[i])
    move_identifier_1_files(I1_files,path_ID,un_identifiers[i])
  }
}

unique_identifiers<-function(filenames){
  #filenames<-get_all_filenames(path)
  vendor.info<-file_info(filenames)
  all_identifiers<-vendor.info$Identifier_1
  #print(all_identifiers)
  num_all_identifiers<-length(all_identifiers)
  #print(num_all_identifiers)
  unique_identifiers<-c()
  for(i in seq(1:num_all_identifiers)){
    curr_identifier<-all_identifiers[i]
    curr_identifier
    if(sum(which(unique_identifiers==curr_identifier))==0){
      unique_identifiers<-c(unique_identifiers,curr_identifier)
    }
  }
  #print(paste("uniqueIDs",unique_identifiers,sep=" "))
  return(unique_identifiers)
}

file_info<-function(files){
  num_files<-length(files)
  msdat<-iso_read_continuous_flow(files[1:num_files])
  file_info<-msdat %>%
    iso_get_file_info(
      select = c(
        #rename?
        Identifier_1 = `Identifier 1`,
        # select columns without renaming
        `Analysis`, `Preparation`,
        # select the time stamp and rename it to `Date & Time`
        Date_and_Time = file_datetime
      ),
      # explicitly allow for file specific rename (for the new ID column)
      file_specific = TRUE #mostly useful with data from different instruments
    )
  # convert from tibble to df
  file_info.df<-as.data.frame(file_info)
}

identifier_1_files<-function(files,identifier_1){
  print(identifier_1)
  identifier_1_files_ind<-pvec(seq_along(files),function(i)
    grepl(identifier_1,files[i],fixed=T))
  print(identifier_1_files_ind)
  # get the indices
  #identifier_1_files_ind<-which(identifier_1_files_ind) #which are TRUE
  # get the identifier_1 file names
  identifier_1_files<-files[identifier_1_files_ind]
  i1.df<-as.data.frame(identifier_1_files)
  colnames(i1.df)<-c(identifier_1)
  return(i1.df)
}
```

Contents of an unsorted directory of .dxf files.
```{r,echo=FALSE}
# print contents of pre-sorted dir
setwd("~/Desktop/EuropaMLMS/rmdMLMS/vignetteData/unsortedData")
list.files()
setwd("~/Desktop/EuropaMLMS/rmdMLMS")
```
<br>

Use *sort_by_identifier_1()* to sort a directory of .dxf files by Identifier_1
```{r,eval=TRUE,message=FALSE,eval=FALSE}
# Use Identifier_1 labels to sort data
unsortedPath<-"~/Desktop/EuropaMLMS/rmdMLMS/vignetteData/sortFolder"
setwd(unsortedPath)
sort_by_identifier_1(unsortedPath)
setwd("~/Desktop/EuropaMLMS/rmdMLMS")
```
<br>
Contents of the sorted directory.
```{r,echo=FALSE}
setwd("~/Desktop/EuropaMLMS/rmdMLMS/vignetteData/sortFolder")
list.files()
setwd("~/Desktop/EuropaMLMS/rmdMLMS")
```
<br>

### Automated quality-control and calibration of a directory of .dxf files

Perform quality checks and remove files that fail any of the checks.

<body>
**Quality checks:**
  <ul>
 1. 
 2. 
 3. 
 4. 
 5. 
 6.
 </ul>
<body>

The list above represents the order of quality checks and a file is removed immediately after it fails a check (before the next check is performed).

**1. Peaks present/number of peaks **<br>
Check that there are more than 0 and fewer than a specified number of peaks.
```{r}
```

**2. Reference peaks**<br>
Check- 
```{r}
```

**3. Reference peaks**<br>
checks-
```{r}
```
<br>

#### Quality-control summary for a directory of .dxf files {.tabset .tabset-fade .tabset-pills}

Output from the quality check and calibration process can summarized in tables and analyzed. 

##### QC1 {.unnumbered}
(tab content)

##### QC2 {.unnumbered}
(tab content)

#### {.unnumbered}
Summaries for analyses that failed checks
<br><br>


#### Internal standards summary for a directory of .dxf files (if standards present)
Internal standards can be checked for a batch of runs (several .dxf files with internal standard experiments).
<br><br>


### Automated separation and quality-control of reference and sample peaks in a combined .csv file for multiple experiments
Processing of weekeqdata (three .csv files) into quality-checked reference and sample peaks 

<br><br>


# UMAP
Illustration of interactive graphics

## Unfiltered data
UMAP for different reactions that have not been quality-checked.<br>
This UMAP uses a directory of unsorted, unfiltered data and is for all the peaks in the data set.

```{r}
unfiltPlot.dat<-read.table("UMAP_NaHCO3_+_NaCl_notQC")
p<-ggplot(unfiltPlot.dat,
       aes(x=x,y=y,color=Identifier_1))+
       geom_point()+
       labs(title="UMAP of NaHCO3 + NaCl (not quality-checked)")
ggplotly(p)
```
<br><br>


## Quality-controlled data
UMAP for different reactions that have been quality-checked.<br>

EXAMPLE of interactive plot with multiple variables

```{r}
library(gapminder) 
p <- gapminder %>%
  filter(year==1977) %>%
  ggplot( aes(gdpPercap, lifeExp, size = pop, color=continent)) +
  geom_point() +
  scale_x_log10() +
  theme_bw()
 
ggplotly(p)
```
<br>


# References


